Flexible Scientific Data Management for Plant Phenomics Research

### 2013-04-12 s4biodiv2013/Rebuttals.txt

 
----------------------- REVIEW 1 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: 3 (strong accept)
REVIEWER'S CONFIDENCE: 5 (expert)
What is your expertise level for the paper?: 5 (excellent) For this paper do you suggest a long presentation (20 min) ?: 1 (yes) For this paper do you suggest a flashtalk (5 min) ?: 3 (maybe) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 3 (maybe) Do you think the paper will open or drive a fruitful discussion during the workshop?: 1 (yes) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 5 (excellent)
ORIGINALITY: from 1 (lowest) to 5 (highest): 5 (excellent) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 5 (excellent)

----------- REVIEW -----------
This paper demonstrate several inovative application of ontologies in Phenomic like : (i) ontology driven data store, (ii) data consistency check, (iii) relation between multiple references ontologies.

For more than ten years, storage and archiving biological data was achieved by first designing a data model representing what was to be stored and then implementing it, generaly in a database. In this paper, we find those two steps but with two major improvment. First, being based on an ontology, the data model is richer that what can be commonly achieved with traditional UML/Database modeling. Secondly, the implementation is not purely based on a database but relies on semantic web technologies like OWL, and are thus very likely to fully preserve the semantic richness of the model.
Furthermore, this datastore is designed for Big Data, and it is of very high interesst to see how ontologies scales and what solutions can be implemented for high volumes of data.

The ontologies are also used for a two level consistency check. There is a core set of rule to ensure the integrability of the data in the system, but it seems to allow the neccessary level of flexibility to ensure data quality without restricting data submission and dataset elaboration.
Furthermore, PODD allows to use several reference ontologies thus easing data interoperability.

This paper shows extensive use of semantic web technologies applied to an important biological field, Phenomics applied to the study of phenotype, genotype and environment relationship in plants. For me, it is therefore perfectly in the scope of S4BioDiv.

It is a very original work, as there no or little phenomic datastore curently relying on ontologies like PODD. Furthermore, it is based on recognized technologies like OWL, RDF and implemented with recognized frameworks like BagIt (cited in the paper) or Hibernate ORM (used in source code). This indicate an excelent technical correcteness.

The experience of PODD is very interresting for every group managing an information system, and this talk should lead to long and fruitfull discussions.

----------------------- REVIEW 2 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: -1 (weak reject)
REVIEWER'S CONFIDENCE: 3 (medium)
What is your expertise level for the paper?: 2 (poor) For this paper do you suggest a long presentation (20 min) ?: 2 (no) For this paper do you suggest a flashtalk (5 min) ?: 2 (no) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 2 (no) Do you think the paper will open or drive a fruitful discussion during the workshop?: 2 (no) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 1 (very poor)
ORIGINALITY: from 1 (lowest) to 5 (highest): 4 (good) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 3 (fair)

----------- REVIEW -----------
The authors describe PODD - an ontology-driven repository to contain data coming from a high-throughput plant phenotyping information.

Though I have few criticisms of the paper itself (and I will raise those in a moment), my main concern is that the authors don't describe how this system is "for biodiversity", and moreover, there is also only passing reference to the semantics of their system in the manuscript.

*****

Rebuttal: We believe that a high level relationship between the semantics of the system and the plant imaging platforms has been described in the paper. We did not set out to describe in intimate details all of the classes and properties that we used to perform this integration in this short workshop paper.

*****

The majority of the manuscript was spent describing neither semantics, nor biodiversity.

*****

Rebuttal: While the majority of the paper describes technical aspects of the system, we believe that its contains direct relationships to both semantics and plant phenomics, the latter of which is a subset of biodiversity.

*****

As such, while the manuscript itself is, in principle, high-quality, it would need to be significantly re-written to be appropriate for this venue.  In an attempt to give the manuscript a more fair review, I attempted to find the connection for myself.  A Google search for HRPPC + biodiversity found several mentions of how the output from the center might be used for biodiversity, but these were extremely generic statements and there didn't appear to be any active project in biodiversity (and if there is, the authors need to describe this in the manuscript).

*****

Rebuttal: The references contain a publication written by HRPPC describing the links between the mesh processing algorithms which generate data which is managed by the PODD system and plant phenomics diversity, which is a subset of biodiversity.

*****

The use of semantics, as described, is difficult to !
 follow.  I understand that they use concepts from, e.g. the Plant Ontology, to annotate their datasets, but that in itself is fairly common and isn't really a publishable achievement.  There is a difficult-to-understand description of the use of OWL as a "rules language", but no clear explanation of why one would use a Description Logic in this way, as opposed to using... for example... a Rules Language.

 *****

 Rebuttal: OWL has been previously used for constraint verification, aka, "Rules", and its approach may be superior to other possible languages. Searching Google for "OWL constraint verifications" returns a range of publications describing its use in this way.

 *****

 As such, the "semantic" aspect of the manuscript isn't particularly compelling either.  The authors didn't discuss if/how semantics is used in their capture and representation of metadata.

 *****

 Rebuttal: Section 2.2 and 2.3 have been updated to include clarifications of the relationship between the (metadata) semantics of the project structure, and the (science) semantics of the links between data, descriptions of processes, results, and conclusions. The project structure semantics are heavily constrained, while the science semantics are less strictly constrained. The science semantics do not attempt to generate high quality upper ontologies, rather they attempt to represent the scientists intentions in a slightly more structured way than otherwise would be possible.

 *****

 The final aspect where semantics was discussed was with respect to "semantic publishing", and that section of the manuscript also wasn't clear.  Presumably by "semantic publishing" the authors mean that it is possible to retrieve fragments of RDF that describe the datasets and their ontology-grounded annotations, but again, this (in and of itself) isn't particularly novel, and the use o!
 f these PODD URIs by other scientists to extend the initial work seems to be the very definition of Linked Data.

 *****

 *****

 I would have been far more supportive of the manuscript if it had described some of their models; explained how the models take advantage of other ontologies (and which); explained which aspects of the data they semantically annotate, why, and how successfully; and how this relates to biodiversity.

The manuscript also feels like it was written by someone extremely familiar with the system, as it uses phraseology that likely only has meaning to them.  For example "External ontologies must currently be mapped into the PODD system to define the expected status of the classes and properties with respect to the basic PODD concepts" - I have read that sentence multiple times and still cannot decypher what it means to "define the expected status of the classes and properties", or what the authors mean by "mapping" external ontologies "into" the PODD system.  Other examples of this include:

"allow scientists to generate datasets before they were certain about the semantic meaning of their work" - what does that mean?

Most of the paragraph beginning "The basic PODD project...", with statements like "annotate images of a plant with a trait in only one case, without annotating other images" and "these horizontal links are not constrained by the basic scientific project structure" (what horizontal links are you referring to, and what are "horizontal links"?)

"any OWL object property that has been defined to link from image acquisition runs to images can be used to define the purpose of an image" - are you saying that images are properties of image acquisition runs?  and that the link between an image acquisition process, and its resulting image, defines the purpose of the image?  How so??

Overall, I thought the manuscript was unnecessarily opaque, didn't emphasise the primary themes of "biodiversity" and "semantics", and over-emphasised many other aspects of their phenotype acquisition and processing workflow.  It's not a bad manuscript, at all, and is likely very well-suited for submission elsewhere (if made slighly less "jargony")!  However, I don't see the fit with this meeting, unless it is extensively re-written.

Best wishes to all the authors!

*Reviewer Name Redacted*

----------------------- REVIEW 3 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: 2 (accept)
REVIEWER'S CONFIDENCE: 3 (medium)
What is your expertise level for the paper?: 4 (good) For this paper do you suggest a long presentation (20 min) ?: 1 (yes) For this paper do you suggest a flashtalk (5 min) ?: 3 (maybe) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 1 (yes) Do you think the paper will open or drive a fruitful discussion during the workshop?: 1 (yes) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 4 (good)
ORIGINALITY: from 1 (lowest) to 5 (highest): 4 (good) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 5 (excellent)

----------- REVIEW -----------
This paper provides an overall view of the Plant Ontology Driven Data repository architecture building to support the High-throughput data generated at the High Resolution Plant Phenomics Centre (Camberra, Australia).
Authors clearly answer major issues of plant Phenomics research applied to the study of phenotype, genotype and environment relationship by providing a flexible and scalable architecture. Moreover the paper shows extensive use of semantic web technologies applied to this field of research. This paper falls completely in the scope of the workshop.

The main contribution of PODD is an OWL ontology which both structure the database model and  check consistency for data. Authors address skillfully the issue of data storage by using robust technologies (e.g. BagIt).

Minor Comment:

I wish appreciate that Authors provide a state of the art for ontology-driven architectures (e.g. iPlant collaborative) and how PODD is distinguised itself from the others.

----------------------- REVIEW 4 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: 2 (accept)
REVIEWER'S CONFIDENCE: 3 (medium)
What is your expertise level for the paper?: 4 (good) For this paper do you suggest a long presentation (20 min) ?: 1 (yes) For this paper do you suggest a flashtalk (5 min) ?: 3 (maybe) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 3 (maybe) Do you think the paper will open or drive a fruitful discussion during the workshop?: 1 (yes) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 4 (good)
ORIGINALITY: from 1 (lowest) to 5 (highest): 4 (good) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 4 (good)

----------- REVIEW -----------
The authors present an intersting approach dedicated to high throughput phenotyping data.
This paper mainly focuses on PODD (Phenomics Ontology Driven Data) properties and the data workflow management. In my opinion, the gap between this paper and state of the art on workflow aspects could be a little more detailled. Overall, the paper is well presented and I believe this work will be relevant for the workshop.

### 2013-04-12 s4biodiv2013/Rebuttals.txt

 *****

 As such, the "semantic" aspect of the manuscript isn't particularly compelling either.  The authors didn't discuss if/how semantics is used in their capture and representation of metadata.

 *****

 Rebuttal: Section 2.2 and 2.3 have been updated to include clarifications of the relationship between the (metadata) semantics of the project structure, and the (science) semantics of the links between data, descriptions of processes, results, and conclusions. The project structure semantics are heavily constrained, while the science semantics are less strictly constrained. The science semantics do not set out to generate high quality upper ontologies, rather they attempt to represent the scientists intentions in a slightly more structured way than otherwise would be possible.

 The capture and representation of metadata is highly platform specific, and while the current reference platform, PlantScan, does not generate scientific semantic metadata, other platforms are free to add semantics to metadata without being biased against their use of metadata semantics.

@@

 *****

 The final aspect where semantics was discussed was with respect to "semantic publishing", and that section of the manuscript also wasn't clear.  Presumably by "semantic publishing" the authors mean that it is possible to retrieve fragments of RDF that describe the datasets and their ontology-grounded annotations, but again, this (in and of itself) isn't particularly novel, and the use o!
 f these PODD URIs by other scientists to extend the initial work seems to be the very definition of Linked Data.

 *****

 Rebuttal: Unfortunately the use of semantic metadata in scientific publications is very limited in practice. Section 6 merely aims to give a quick description of how PODD enables the long-term stability of projects so they can be used as references in static peer reviewed publications, while ontologies are simultaneously updated for use in other projects. This enables scientists to have more control than a raw Linked Data platform where there are no guarantees relating to the stability of the system in general.

@@

 *****

 I would have been far more supportive of the manuscript if it had described some of their models; explained how the models take advantage of other ontologies (and which); explained which aspects of the data they semantically annotate, why, and how successfully; and how this relates to biodiversity.

 Rebuttal: There are no measures for the successful use of external ontologies, as the definitions of external ontologies are always in part defined by the community based on both discussions and based on the way they are used. The set of ontologies that are currently in use is described in Figure 2.

@@

 Rebuttal : TODO: Fix this

 *****

Overall, I thought the manuscript was unnecessarily opaque, didn't emphasise the primary themes of "biodiversity" and "semantics", and over-emphasised many other aspects of their phenotype acquisition and processing workflow.  It's not a bad manuscript, at all, and is likely very well-suited for submission elsewhere (if made slighly less "jargony")!  However, I don't see the fit with this meeting, unless it is extensively re-written.

 Rebuttal : We feel that the reviewer did not recognise the attempt to provide a high level view of the integrated PODD and PlantScan system and how it is being used by HRPPC.

 The reviewer requests that the manuscript use less jargon, but also detail more specifically what ontologies are in use and how they are being practically used.

 We broadly agree with the comments and have made changes to clarify various points.

### 2013-04-12 s4biodiv2013/PODD-S4BIODIV-Submission.tex
make changes based on reviewer comments

The basic PODD project is structured as a tree, which, when used with
PlantScan\textsuperscript{\texttrademark}, represents the different components
of a scientific project as
top-level branches. These include a branch for raw data, along with separate
branches for results, analysis, and publications related to the project. In the
case of raw data, the semantics are not clear and cannot be defined by the
platforms collecting the data. 
% Scientists are able to sparsely attach meaning
% using horizontal links between different branches, using the OWL Open World
% Assumption to allow extra semantic annotations to be attached in future. 
For example, they may annotate images of a plant with a trait in only one case,
without annotating other images. These horizontal links are not constrained by
the basic scientific project structure, so they are able to entail novel
meanings independent of the ontology, including untested hypotheses, process
annotations and meaningful scientific results.

\subsection{Semantic validation}
PODD validates experiments using independently configurable rules based on OWL
(Web Ontology Language) schemas. Although PODD currently only supports OWL as a
rules language, it could be easily extended in other cases to use different
systems such as N3, RDFS, SPARQL, or SPIN as rules languages \cite{Fuerber2010}.

OWL is used to determine whether projects are both internally consistent, with
all objects having an explicit RDF type, and whether they are consistent with
the ontologies that they import. The main point that PODD focuses on for
validation is determining that mappings are consistent with the basic PODD
predicates and classes, so that they can be used to render both static HTML
pages and provide HTML based editing facilities. For example, any OWL object
property that has been defined to link from image acquisition runs to images can
be used to define the purpose of an image. This may include any combination of
the defined properties, without requiring all of the properties to be used.
% The parent object types and properties used by PODD to establish basic ground
% rules for rendering and validation are defined in a small base schema ontology.
General scientific properties and phenotype specific properties are defined
in larger extension ontologies as illustrated in Figure~\ref{poddontologies}.

>>>>

In the case of PlantScan\textsuperscript{\texttrademark} a PODD project contains top level branches describing the 
various parts of a scientific project. These include a branch for raw data, along with separate
branches for results, analysis, and publications related to the project. In the
case of raw data, the semantics are not necessarily clear and are not easily defined by the
automated platforms collecting the data. 
% Scientists are able to sparsely attach meaning
% using horizontal links between different branches, using the OWL Open World
% Assumption to allow extra semantic annotations to be attached in future. 
The scientist may then semantically link the data with results, conclusions, and external ontologies.
For example, a scientist may annotate the data objects representing images of a plant with a trait 
that is defined in an external community ontology. They may also annotate the image with a link 
to a trait that is defined inside of the project, particularly in cases where the trait is novel 
and not represented in a community ontology.
%These horizontal links are not constrained by
%the basic scientific project structure, so they are able to entail novel
%meanings independent of the ontology, including untested hypotheses, process
%annotations and meaningful scientific results.

\subsection{Semantic validation}
PODD validates scientific project descriptions using independently configurable constraints based on OWL
(Web Ontology Language) ontologies. Although PODD currently solely supports OWL for 
constraint verification, it could be easily extended in other cases to use different
systems such as N3, RDFS, SPARQL, or SPIN as rules languages \cite{Fuerber2010}.

OWL is used to determine whether projects are both internally consistent, with
all objects having an explicit RDF type, and whether they are consistent with
the ontologies that they import. 
%The main point that PODD focuses on for
%validation is determining that mappings are consistent with the basic PODD
%predicates and classes, so that they can be used to render both static HTML
%pages and provide HTML based editing facilities. 
For example, any OWL object
property that has been defined to link from image acquisition runs to images can
be used to define the purpose of an image. This may include any combination of
the defined properties, without requiring all of the properties to be used.
% The parent object types and properties used by PODD to establish basic ground
% rules for rendering and validation are defined in a small base schema ontology.
General scientific properties and phenotype specific properties are defined
in optional extension ontologies as illustrated in Figure~\ref{poddontologies}.
These are used by scientists to annotate their projects with concepts specific 
to their field, without requiring other scientists using the system to use phenotype 
properties to annotate their projects.
